---
title: "Stat_rethinking_Ch3"
author: "Jacob Nesslage"
date: "2024-01-31"
output: html_document
---
# Basic Procedures for Manipulating Posterior Distributions


## Homework 

The easy problems use the samples from the posterior distribution for the globe tossing examples. THis code will give you a specific set of samples, so that you can check your work.

```{r generate posterior distribution and select samples}
p_grid <- seq(from=0,to=1,length.out=1000)
prior <- rep(1,1000)
likelihood <- dbinom(6,size=9,prob=p_grid)
posterior <- likelihood * prior
posterior <- posterior/sum(posterior)

set.seed(100)
samples <- sample(p_grid,prob=posterior,size=1e4,replace=T)
```

And, a visualization

```{r,warn=F}
require(rethinking)
dens(prior)
dens(likelihood)
dens(posterior)
dens(samples)
```


### 3E1: How much posterior probability lies below 0.2? (use samples)
```{r}
sum(samples < 0.2)/1e4
```

### 3E3: How much posterior probability lies between 0.2 and 0.8?
```{r}
sum(samples > 0.2 & samples < 0.8)/1e4
```


### 3E5: 20% of the posterior probability lies above which value of p?
```{r}
quantile(samples,0.2)
```

### 3M2
```{r}
p_grid <- seq(from=0,to=1,length.out=1000)
prior <- rep(1,1000)
likelihood <- dbinom(8,size=15,prob=p_grid)
posterior <- likelihood * prior
posterior <- posterior/sum(posterior)

set.seed(100)
samples <- sample(p_grid,prob=posterior,size=1e4,replace=T)

HPDI(samples,prob=0.9)
```
Sanity check!

```{r}
dens(samples,show.HPDI=0.9)
```


### 3M6
```{r}
#Create empty data frame for simulation
sim <- data.frame(count=double(),width=double())

#Try a few different sample sizes by looping
for (i in seq(1000,3000,by=100)){
  
  #Set up grid approximation as normal
  p_grid <- seq(from=0,to=1,length.out=1000)
  prior <- rep(1,1000) #uniform prior
  
  #Here, we want to model the probability of getting 70% water in our likelihood
  #Using a randomly generated value here seems fair
  obs <- rbinom(1,i,prob=0.7)
  
  #Now, generate the likelihood for a given sample size
  likelihood <- dbinom(obs,size=i,prob=p_grid)
  
  #Posterior calculated as normal
  posterior <- likelihood * prior
  posterior <- posterior/sum(posterior)
  
  #Take samples from the posterior for calculation of a percentile interval
  samples <- sample(p_grid,posterior,size=1e4,replace = T)
  interval <- PI(samples,prob=0.99)
  
  #Calculate 99% percentile interval width
  width <- interval[2]-interval[1]
  
  #Add count and 99% percentile interval width to sim results
  sim_result <- data.frame(i,width)
  sim <- rbind(sim,sim_result)
}
```

```{r}
require(tidyverse)
ggplot(sim, aes(x = i, y = width)) +
  geom_point() +
  geom_line() +
  labs(x = "Tosses", y = "Interval Width") +
  theme(panel.spacing.x = unit(2, "lines")) +
  theme_bw()
```


### 3H3
```{r}
#Set up grid approximation as normal
  p_grid <- seq(from=0,to=1,length.out=1000)
  prior <- rep(1,1000) #uniform prior
  
  #Here, we want to model the probability of getting 70% water in our likelihood
  #Using a randomly generated value here seems fair
  sim_births <- rbinom(10000,200,prob=0.5)/200
  #sim_births <- sim_births
  
  #Now, generate the likelihood for a given sample size
  likelihood <- dbinom(111,size=200,prob=p_grid)
  
  #Posterior calculated as normal
  posterior <- likelihood * prior
  posterior <- posterior/sum(posterior)
  
  #Take samples from the posterior for calculation of a percentile interval
  samples_actual <- sample(p_grid,posterior,size=1e4,replace = T)
  #samples_sim <- sample(p_grid,sim_births,size=1e4,replace = T)
```

```{r}
dens(sim_births,show.HPDI=0.95)
dens(samples_actual,add=T,show.HPDI=0.95)
```

