---
title: "Stat_rethinking_Ch3"
author: "Jacob Nesslage"
date: "2024-01-31"
output: html_document
---
# Basic Procedures for Manipulating Posterior Distributions


## Homework 

The easy problems use the samples from the posterior distribution for the globe tossing examples. THis code will give you a specific set of samples, so that you can check your work.

```{r generate posterior distribution and select samples}
p_grid <- seq(from=0,to=1,length.out=1000)
prior <- rep(1,1000)
likelihood <- dbinom(6,size=9,prob=p_grid)
posterior <- likelihood * prior
posterior <- posterior/sum(posterior)

set.seed(100)
samples <- sample(p_grid,prob=posterior,size=1e4,replace=T)
```

And, a visualization

```{r,warn=F}
require(rethinking)
dens(prior)
dens(likelihood)
dens(posterior)
dens(samples)
```


### 3E1: How much posterior probability lies below 0.2? (use samples)
```{r}
sum(samples < 0.2)/1e4
```

### 3E3: How much posterior probability lies between 0.2 and 0.8?
```{r}
sum(samples > 0.2 & samples < 0.8)/1e4
```


### 3E5: 20% of the posterior probability lies above which value of p?
```{r}
quantile(samples,0.2)
```

### 3M2
```{r}
p_grid <- seq(from=0,to=1,length.out=1000)
prior <- rep(1,1000)
likelihood <- dbinom(8,size=15,prob=p_grid)
posterior <- likelihood * prior
posterior <- posterior/sum(posterior)

set.seed(100)
samples <- sample(p_grid,prob=posterior,size=1e4,replace=T)

HPDI(samples,prob=0.9)
```
Sanity check!

```{r}
dens(samples,show.HPDI=0.9)
```


### 3M6
```{r}

```

### 3H3
```{r}

```

